---
title: "Week 4 - Assignment - Lahman Teams Baseball Data"
author: "Mohamed Hussein"
date: "7/16/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction to the Shiny App

In this assigment, I use baseball teams historic data from the R package {Lahman} to visualize how run differential can be used to predict a team's winning percentage that is:

- `(Runs - Runs allowed)`
- `( Wins / Wins + Losses)`

## Introduction to Data

Here is a snapshort of the data - which in its current entirety contains 2925 observations and 48 variables. Fore the purpose of this simple app, I use only records from the year 2000 to 2019.

```{r, comment=""}
library(Lahman)
data("Teams")
head(Teams[Teams$yearID > 2000, c(1:8)])
```

## Main Plot - Win % vs Run Differential
In the app, you are able to choose different dates and see changes in the slope and intercept 
```{r echo=FALSE}
library(ggplot2)
Teams$RD <- with(Teams, R - RA)
Teams$Wpct <- with (Teams, W / (W + L))
linfit <- lm(Wpct ~ RD, data = Teams)

ggplot(data=Teams, aes(x=RD, y=Wpct)) + 
            geom_point(alpha=0.3) + 
            geom_abline(slope = linfit$coefficients[2], intercept = linfit$coefficients[1]) +
            labs(title = "Plot of a Team's Run Differential vs Win Percentage", 
                 subtitle = paste0("Slope: ", round(linfit$coefficients[1],5), 
                                  ", Intercept: ", round(linfit$coefficients[2],5))) +
            xlab("Run Differential") + ylab("Winning Percentage") +
            theme(plot.title = element_text(face="bold"))
```

## Plot of Residuals

```{r cars, echo = FALSE}
#prediction and Residuals
Teams <- Teams[Teams$yearID >= 2000,]
Teams$RD <- with(Teams, R - RA)
Teams$Wpct <- with (Teams, W / (W + L))
linfit <- lm(Wpct ~ RD, data = Teams)
Teams$linWpct <- predict(linfit)
Teams$linResids <- residuals(linfit)

#Graph of Residuals
plot(Teams$RD, Teams$linResids,
     xlab="Run Differential",
     ylab="Residual")
abline(h=0, lty = 3)
```

## Thank you

Thank you for taking the time to review my work. Appreciated!

